---
title: "KobeShots"
output: html_notebook
---


In this report we are going to analyse a dataset containing a selection of Kobe Bryant's shots throughout his career. The goal of the analysis is to build a binary classification model that is able to predict the outcome of the shot given some informations about the shots, referred to as *features* or *predictors* from now on.

The dataset is publicy available in the CSV format and contains ca. 30K shots, described by means of 25 attributes.\newline
In the first section we are going to have a look at those attributes and perform a cleaning and transformation phase in order to make the dataset more suitable for further analysis. In the rest of the report we will use different techniques in the attempt of building a machine learning model that is able to correctly predict the \textit{shot\_made\_flag} attribute, which will be also called *target variable*.\newline
The analysis will be carried out using R 3.6 as well as different libraries, which will be pointed out each time.

### 1. Data exploration

The first phase in the analysis of a dataset must always be an exploration of the data provided. This may include a cleaning phase, in case some values or attributes are not needed, and a feature engineering phase, in which new predictors are created basing on the existing attributes and some kind of domain knowledge.\newline
After loading the dataset in the R environment, we proceed with a coarse data exploration and cleaning. The dataset contains 30.697 records, however many of them are missing some values and therefore we decide to prune them: this leaves us with 25.697 records.

```{r}
kobe <- read.csv(file = 'data.csv')
kobe=na.omit(kobe)
nrow(kobe)
attach(kobe)
#summmary(kobe) # have a look at the dataset
```


Subsequently we have a look at the attributes provided in the dataset. We notice that some features are redundant: for example, *team_name* and *team_id* take on a single value, while *lat* and *lon* provide the same information as *loc_x* and *loc_y* on a different scale.\
It can also be useful to plot some of the attributes, in order to gain an intuitive feeling for the data we are dealing with. For example, we can plot loc_x and loc_y marking the successful shots in blue and the unsuccessful ones in red.


```{r}
dotsCol = rep("red", length(shot_made_flag))
dotsCol[which(shot_made_flag==TRUE)] = "blue"
plot(loc_y, loc_x, type="p", cex=0.1, pch=20, col=dotsCol)
```

#### 1.2 FAMD
```{r}
library("FactoMineR")
library("factoextra")
famd = FAMD(kobe, graph = FALSE)
get_eigenvalue(famd)
fviz_screeplot(famd)
var = get_famd_var(famd)
var$contrib
fviz_famd_var(famd, repel = TRUE)
```

### 2. Feature engineering
Using some domain knowledge - e.g. that an NBA court is 94 by 50 feet - we can imply that loc_x and loc_y are expressed in tenth (1/10 of a foot). Since many of the analysis we are going to present base on some kind of linearity between the predictors and the target variable we prefer not to have a big differentiation between similar shots, such as those taken from the right angle and from the left angle, which have comparable rate of success.\
Therefore we add two new features, called *feet_x* and *feet_y*, which express the absolute value of the original attributes, divided by 10 (i.e. converted into feet). We create also an additional value, *shot_side*, which signals whether the shot was taken from the right side (i.e. in the positive range of loc_x) or from the left side. In this way it will be easier to point out in the following sections if the side does play a role in the outcome of the shot or not.

Another example of feature engineering is the use of different coordinate systems to express the shot location. Along with the cartesian system used in feet_x and feet_y, here we want to also employ the polar coordinates, which consider the shot distance and the angle between the shooter and the rim. In fact, even though there is a one-to-one mapping between the two systems, this is not achieved through a linear function: for this reason one of them could e.g. better fit the linear relationship in the log-odds of the prediction, which would reduce the bias of the logistic regression.\
We conclude the first phase doing the transformation of one more attribute: specifically, we consider the attribute *season*, which is qualitative in the original dataset. Since we expect to see a gradual evolution in Bryant's shots over his career, we rather prefer to consider this variable as numerical, so that we the model can integrate more easily the temporal informations in its prediction.


```{r}
feet_x = abs(loc_x)/10
feet_y = loc_y/10
kobe$shot_distance = sqrt(feet_x^2 + feet_y^2) # distance in the polar coordinates
shot_angle = atan2(feet_y, feet_x)             # angle in the polar coordinates
shot_side = ifelse(loc_x >= 0, "Right", "Left")
kobe = cbind(kobe, shot_angle, feet_x, feet_y, shot_side)
season = substr(season, 1, 4)
season = strtoi(season)
kobe$season = season
kobe = kobe[,-which(names(kobe) %in% c("minutes_remaining","seconds_remaining","team_name", "team_id", "matchup", "shot_zone_range", "game_event_id", "game_id", "game_date", "shot_id", "lat", "lon", "loc_x", "loc_y"))]
```


```{r}
rareVars = names(which(summary(action_type) < 50))
rareIndex = action_type %in% rareVars
kobe$action_type = factor(kobe$action_type, levels=c(levels(action_type), "Other"))
kobe$action_type[rareIndex] = "Other"
action_type = kobe$action_type
```

### 2. Logistic regression

#### 2.1 First model
We start by building a first, naive logistic regression model. First, we have which predictors we use to train our model: at this points, it's important not to include correlated attributes in the predictors. It is important to remove strong correlations among the attributes since they can negative impact: in particular, they can increase the variance of the coefficients and hence harm the reliability of the result.\
We can check the correlation among continuous variables building a correlation matrix:

```{r}
#cor(kobe[,-c(1,7,8,9,14)])
```

Here we see that most of the variables are not strongly correlated. This does not hold for feet_x, feet_y, shot_distance and shot_angle. This is no wonder, since shot_distance and shot_angle have been created starting from the other two.\
Similarly, *combined_shot_sype* is a generalization of *shot_type*, while *shot_zone_area*, *shot_zone_basic* and *shot_zone_range* are a generalization of the shot coordinates, therefore we ignore them for now.

```{r}
logreg = glm(shot_made_flag~.-combined_shot_type-shot_zone_area-shot_zone_basic-feet_x-feet_y , family=binomial, data=kobe)
summary(logreg)
```

We can see that in order to build the model all categorical predictors have to be transformed into dummy variables. In the case of binary predictors, such as shot_side, this simply means marking one value (in this case "Right") with 1 and the other one with 0. In the case of larger domains, there has to be one dummy variable for every possible value (minus one), which is 1 if the predictor takes on that value. Accordingly, the related coefficient indicates how the output changes when the attribute takes on that value.\
We can have an insight on the statistical significance of our prediction by looking at the z-value and the corresponding p-value (displayed in the script as *Pr(>|z|)*). The z-value is a score that says how far each coefficient is from 0, which in turn would imply there is no relationship between the predictor and the response. The z-value is computed using the sample standard error: we can then look at the p-value to consider how likely is that the coefficient is 0, given its z-value.\
In this first model, we can see that the dummy variables derived from opponent have quite high p-values, which suggests that the outcome of a shot does not depend on the opposing team. On the contrary, the coefficient of shot_angle is very unlikely to be 0, even though it has a small absolute value (0.12): that is because its standard error is so little (0.03) that the z-score is cranked up, so as we can rely on the statistical significance of the coefficient. If we consider how shot_angle was constructed, we see that its value is 0 for shots taken from the base line and $\pi$/2 for shots taken face the rim. Therefore, the value of 0.12 means that shots taken in front of the rim have higher odds of being successful, since they have higher values of shot_angle.\

#### 2.2 Evaluation

##### 2.2.1 Validation set
The simplest way to evaluate a model is to use validation set. This approach consists in setting aside a part of the samples during the construction of the model and using them only in the evaluation. Common size proportions between training and validation set are 70%-30% and 80%-20%.\
Clearly, in order to evaluate the models from the last section we have to repeat also the training phase, because we are not allowed to use the samples in the validation phase for building the model. However, we expect that the results obtained sofar also hold for the new training set, since its samples are chosen randomly.\
Doing so, we can 

```{r}
set.seed(0)
# Build training and validation set
train_id = sample(nrow(kobe), 0.7*nrow(kobe))
test_x = kobe[-train_id,]
test_y = shot_made_flag[-train_id]
# Rebuild logistic regression models
logreg1 = glm(shot_made_flag~.-combined_shot_type-shot_zone_area-shot_zone_basic-feet_x-feet_y , family=binomial, data=kobe, subset=train_id)
# Evaluate models
probs1 = predict(logreg1, test_x, type="response")
preds1 = ifelse(probs1>.45, 1, 0)
acc1 = sum(preds1==test_y)/length(test_y)
```


##### 2.2.2 Cross-validation
```{r}
# TODO: review..is it correct?
library(boot)
costFun <- function(r, pi) mean(abs(r-pi)> 0.5)
logreg1 = glm(shot_made_flag~.-shot_zone_area-shot_zone_basic-feet_x-feet_y , family=binomial, data=kobe)
cv1 = cv.glm(kobe, logreg1, cost=costFun, K=10)$delta[1]
logreg2 = glm(shot_made_flag~.-shot_zone_area-shot_zone_basic-feet_x-feet_y-playoffs-season-opponent-shot_side, family=binomial, data=kobe)
cv2 = cv.glm(kobe, logreg2, cost=costFun, K=10)$delta[1]
```

#### 2.3 Improving the model
The presence of insignificant predictors can negatively impact on the model, for example by increasing its variance. Therefore we now refine our first model by ignoring such predictors; in the next section we will evaluate both models and see whether this decision leads to a better performance.

```{r}
# 1. ignore attributes with high p-value
logreg2 = glm(shot_made_flag~.-combined_shot_type-shot_zone_area-shot_zone_basic-feet_x-feet_y-playoffs-season-shot_distance-opponent-shot_side, family=binomial, data=kobe, subset=train_id)
probs2 = predict(logreg2, test_x, type="response")
preds2 = ifelse(probs2>.45, 1, 0)
acc2 = sum(preds2==test_y)/length(test_y)
```

```{r}
# 2. cartesian coordinates
logreg3 = glm(shot_made_flag~.-combined_shot_type-shot_zone_area-shot_zone_basic-shot_angle-shot_distance-playoffs-season-shot_side, family=binomial, data=kobe, subset=train_id)
probs3 = predict(logreg3, test_x, type="response")
preds3 = ifelse(probs3>.45, 1, 0)
acc3 = sum(preds3==test_y)/length(test_y)
```

```{r}
# model variance
logreg4 = glm(shot_made_flag~.-action_type-shot_zone_area-shot_zone_basic-feet_x-feet_y , family=binomial, data=kobe, subset=train_id)
coef(summary(logreg1))[,4][27:31]
coef(summary(logreg4))[,4][7:11]
probs4 = predict(logreg4, test_x, type="response")
preds4 = ifelse(probs4>.45, 1, 0)
acc4 = sum(preds4==test_y)/length(test_y)
```

#### 2.4 Model selection


##### 2.4.1 Forward stepwise selection
<!--- Tell you use R^2 --->
```{r}
library(leaps)
names(kobe)
logreg.fwd = regsubsets(shot_made_flag~action_type+combined_shot_type+period+playoffs+season+shot_distance+shot_type+shot_zone_area+shot_zone_basic+opponent+shot_angle+feet_x+feet_y+shot_side, data=kobe[train_id,], nvmax=25, method="exhaustive")
kobe.fw = kobe
kobe.fw = kobe.fw[,-c(1,11)]
logreg.fwd = regsubsets(shot_made_flag~., data=kobe.fw[train_id,], nvmax=25, method="forward")
coef(logreg.fwd, 5)
```

```{r}
test.mat = model.matrix(shot_made_flag~., data=kobe.fw[-train_id,])
val.errors = rep(NA, 25)
for (i in 1:25) {
  coefi = coef(logreg.fwd, id=i)
  preds = test.mat
}
```

##### 2.4.2 Backward stepwise selection
```{r}
```

##### 2.4.3 Pca
```{r}
pr.out = prcomp(kobe, scale=TRUE)
```

### 3. Decision tree

### 3.1 Bagging

### 3.1 Random forest

### 4. Lda

### 5. Svm

### 6. Knn

### (7. Evaluation metrics)
<!--
hyperparameters (logistic regression threshold)
other metrics (precision, recall, roc curve)
-->